---
title: "Final Project Code"
author: "Megan Mason, Gwen Hammond, and Ryan Jensen"
date: "2025-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## DS 2020 Final Project: Do American Lifestyles correlate with how you like your steak?

## The Data Set

-   This data set looks at factors and risky behaviors that are found in the American lifestyle. The purpose of this data set is to predict how lifestyle choices correlate with how Americans like there steak. The data set contains 551 observations and has 15 variables which are mainly categorical.

## Cleaning the Data

-   Megan started off cleaning the data by re-names the columns so that they are easier to work with.

-   Gwen created a subset of the data where missing values are excluded.

-   Ryan created subsets of people who do eat steak and who don't eat steak.

```{r}
#Megan's Code 
library(readr)
steak <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/refs/heads/master/steak-survey/steak-risk-survey.csv", stringsAsFactors = FALSE)

names(steak)[names(steak) == "Consider.the.following.hypothetical.situations...br.In.Lottery.A..you.have.a.50..chance.of.success..with.a.payout.of..100...br.In.Lottery.B..you.have.a.90..chance.of.success..with.a.payout.of..20...br..br.Assuming.you.have..10.to.bet..would.you.play.Lottery.A.or.Lottery.B."] <- "Lottery"
names(steak)[names(steak) == "Do.you.ever.smoke.cigarettes."] <- "Smoke.Cigarettes"
names(steak)[names(steak) == "Do.you.ever.drink.alcohol."] <- "Drink.Alcohol"
names(steak)[names(steak) == "Do.you.ever.gamble."] <- "Gamble"
names(steak)[names(steak) == "Have.you.ever.been.skydiving."] <- "Skydiving"
names(steak)[names(steak) == "Do.you.ever.drive.above.the.speed.limit."] <- "Speed"
names(steak)[names(steak) == "Have.you.ever.cheated.on.your.significant.other."] <- "Affair"
names(steak)[names(steak) == "How.do.you.like.your.steak.prepared."] <- "Prepared"
names(steak)[names(steak) == "Do.you.eat.steak."] <- "Consume.Steak"
names(steak)[names(steak) == "Location..Census.Region."] <- "Location"


View(steak)

```

```{r}
#Gwen's Code
clean_steak <- steak[complete.cases(steak), ]

anyNA(clean_steak)


```

